Bee brain size is positively associated with urban lifestyles but not with positive population trends

### 2022-11-23 Manuscript/Ms.Rmd
Edits

Our dataset contains measurements of brain and body size for bee specimens captured by hand netting in different areas of the East Coast of United States and Europe (i.e., Spain and Netherlands). In total, the dataset includes `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` different families and `r nlevels(factor(unique(tax$genus)))` genera. We only considered female specimens because most bees exhibit sexual dimorphism and behavioral differences between sexes that can be correlated with the size and structure of the brain [@streinzer2013; @roselino2015]. In addition, because brain size can change with age and experience of the different individuals within a species [@durst1994; @withers2008], we limited our sampling to specimens foraging on flowers. Note that 93 of the species used in our study were previously published with their respective trait measurements in @sayol2020.

@@

## (d) Species occurrences data

Global Biodiversity Information Facility (GBIF, <http://www.gbif.org/>)

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure 1.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species (N = 85). A) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark. B) East Coast of Unites States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland.", fig.pos="p"}

@@

p2 = ggplot(Europe) +
geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
theme(panel.background = element_rect(fill = "white"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
ylab("Latitude") +
xlab("Longitude") +
ggtitle("A") +
theme(legend.position = c(0, 1),legend.justification = c(0, 1),
legend.background = element_blank(),
legend.box.background = element_blank(),
legend.key = element_blank(), 
legend.title = element_text(face = "bold"),
axis.title = element_text(face="bold", size = 14),
plot.title = element_text(face="bold", size = 18),
legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
scale_y_continuous(breaks = seq(40, 60, by = 6)) +
scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
guides(fill = guide_colourbar(title.position = "top", title.hjust = .5)) +
ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15)

@@

p2 = ggplot(canada) + 
geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2) + 
geom_sf(data= usa, fill="gray80",color = "black", alpha = 1,size=0.2) +
geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
coord_sf(xlim = c(-82, -67), ylim = c(33, 47))  +
theme(panel.background = element_rect(fill = "white"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
ylab("Latitude") +
xlab("Longitude") +
ggtitle("B") +
theme(legend.position = c(0, 1),legend.justification = c(0, 1),
legend.background = element_blank(),
legend.box.background = element_blank(),
legend.key = element_blank(),
legend.title = element_text(face = "bold"),
axis.title = element_text(face="bold", size = 14),
plot.title = element_text(face="bold", size = 18),
legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
scale_y_continuous(breaks = seq(33, 47, by = 4)) +
scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))+
ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))

@@

library(patchwork)
panel1 + panel2

@@

@@

**Animals**

Most bees were opportunistically captured by hand netting in different regions of New York State (USA), and complemented with bees captured from Spain and the Netherlands for species with a holarctic distribution. We captured 145 individuals from 44 species, kept them in cold storage and sent them to the laboratory to be processed. Once the bees were in the laboratory, they were anesthetized in cold (Crook, 2013) and decapitated. Bees were identified by expert taxonomists (Parker Gambino and Ivo Raemakers). The species used for the analysis can be consulted in Table S2. We measured the inter-tegular distance, as it is a good proxy of body size (Kendall et al., 2019). To prevent degradation, the head was fixed in 4% paraformaldehyde with phosphate buffer saline (PBS). Brains were extracted from the head capsule, and tracheae and fat bodies were removed to maximize accuracy in neural tissue weight. Brains extracted were placed on a small piece of tared Parafilm® and exceeding fixative solution was dried from the brain using Kimwipes® tissues. Finally the brain was weighted in a microbalance. Brain weight was used as a proxy of brain size (Sayol et al., 2019 in preparation).

>>>>

Our dataset contains measurements of brain and body size for bee specimens captured by hand netting in different areas of the East Coast of United States and Europe (i.e., Spain and Netherlands). In total, the dataset includes `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` different families and `r nlevels(factor(unique(tax$genus)))` genera. We only considered female specimens because most bees exhibit sexual dimorphism and behavioral differences between female and male individuals [@shreeves2008] that often correlate with the size and structure of the brain [@streinzer2013; @roselino2015]. In addition, because brain size can change with age and experience of the different individuals within a species [@durst1994; @withers2008], we limited our sampling to specimens foraging on flowers. Note that 93 of the species used in our study were previously published with their respective trait measurements in @sayol2020.

@@

## (d) Occurrence data

#Calculate study site areas
#Load libraries
library(sf)
library(sfheaders)

#EUROPE:
#Read data
euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
#Add area col
euro_map$area <- st_area(euro_map)
#Create dataframe with the correct units
d1 = euro_map %>% group_by(CNTR_CODE) %>% summarize(st_union(geometry), area_NAME = sum(area)) %>% 
mutate(areakm2 = units::set_units(area_NAME, value = "km^2"))
#Calculate area
eukm2 = prettyNum(as.numeric(round(sum(d1$areakm2),0)),big.mark=",")

#USA:
#Read data
usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
#Not an sf object, convert into multypolygon
usa = sfheaders::sf_multipolygon(
  obj = usa_states_1
  , multipolygon_id = "group"
  , x = "long"
  , y = "lat"
) %>% st_set_crs(4326)
#Calculate area
usa$area <- st_area(usa)
#Create dataframe with the correct units
d2 = usa %>% group_by(group) %>% summarize(st_union(geometry), area_NAME = sum(area)) %>% 
mutate(areakm2 = units::set_units(area_NAME, value = "km^2"))
#Sum areas
usakm2 = prettyNum(as.numeric(round(sum(d2$areakm2),0)),big.mark=",")

We downloaded occurrence information for all species from the Global Biodiversity Information Facility (GBIF; http://www.gbif.org/) for North America and Europe. The data was downloaded through R programming language with the help of the function *occ_download* from the package *rgbif* version 3.7.3 [@chamberlain2017]. The occurrence data can be downloaded from https://doi.org/10.15468/dl.nenuem where just records for the countries of interest were retrieved. Because the records were biased towards some areas that were associated with political boundaries, we selected the states or countries with highest density of records for our set of species. For North America, we selected states located on the East Coast of United States (Figure 1A) that covered an approximate area of `r usakm2` km$^{2}$. These states were: Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland. Similarly, we selected countries located on the north and center of Europe (Figure 1B) that represented a total area of `r eukm2` km$^{2}$. These countries were: England, Belgium, Netherlands, Germany and Denmark.

External data information [...] Origin?

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure 1.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species (N = 85). A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland. B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.", fig.pos="p"}

@@

p2 = ggplot(Europe) +
geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
theme(panel.background = element_rect(fill = "white"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
ylab("Latitude") +
xlab("Longitude") +
ggtitle("B") +
theme(legend.position = c(0, 1),legend.justification = c(0, 1),
legend.background = element_blank(),
legend.box.background = element_blank(),
legend.key = element_blank(), 
legend.title = element_text(face = "bold"),
axis.title = element_text(face="bold", size = 14),
plot.title = element_text(face="bold", size = 18),
legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
scale_y_continuous(breaks = seq(40, 60, by = 6)) +
scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
guides(fill = guide_colourbar(title.position = "top", title.hjust = .5)) +
ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15)

@@

p2 = ggplot(canada) + 
geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2) + 
geom_sf(data= usa, fill="gray80",color = "black", alpha = 1,size=0.2) +
geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
coord_sf(xlim = c(-82, -67), ylim = c(33, 47))  +
theme(panel.background = element_rect(fill = "white"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
ylab("Latitude") +
xlab("Longitude") +
ggtitle("A") +
theme(legend.position = c(0, 1),legend.justification = c(0, 1),
legend.background = element_blank(),
legend.box.background = element_blank(),
legend.key = element_blank(),
legend.title = element_text(face = "bold"),
axis.title = element_text(face="bold", size = 14),
plot.title = element_text(face="bold", size = 18),
legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
scale_y_continuous(breaks = seq(33, 47, by = 4)) +
scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))+ 
ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))

@@

library(patchwork)
panel2  + panel1

@@

## (e) Construction of the phylogenetic tree

@@
